---
title: "Seagle CH 3-4 Test"
author: "Adam Seagle"
date: "4/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(readxl)
library(GGally)
```
1)
```{r}
CF <- read.csv("CFB2018completeISLR.csv")
```

```{r}
library(tidyverse)

cfb <- CF %>% select(4,6,7,9,12,13,15,18,19,21,24,25,27,30,31,33)
```

```{r}
cor(cfb)
```

Based on the correlation matrix values above the single best predictor of Zsagarin, based on the potential predictors, is the Fravg predictor. I can tell this because Fravg has the hoghest correlation with the Zsagarin variable with a correlation of 0.568.

```{r}
CFmr <- lm(Zsagarin ~ Fravg + Soavg + Jravg, data = CF)
summary(CFmr)
```

```{r}
CFmr2 <- lm(Zsagarin ~ Fravg + Soavg + Jravg + conference, data = CF)
summary(CFmr2)
```
Once I added the conference predictor to my model, I found that my R-squared value went up slightly but my F-statistic went down quite a bit. I believe this happened due to only one of the conferences being statistically significant while the others have very high p-values. I also saw that this left two of my three original predictors (Fravg and Jravg) with roughly the same p-value but decreases the p-value of my Soavg predictor. 

```{r}
library(car)

vif(CFmr2)
```
For my model the Variance Inflation Factors are all below 5 so they are not concerning to me when I look at them. This tells me there is not siginificant multicollinearity within my model.

```{r}
plot(CFmr2)
```
I like the way that the above plots look for the most part. However, in the Residuals vs Fitted, Scale-Location and in the Residuals vs Leverage Plots I do see a slight decrease at the beginning but then it levels out. I would like to see if I could make it a straight line all the way across not just from the halfway point going forward. 

```{r}

CFmr3 <- lm(Zsagarin ~ Fravg + Jravg + Soavg:coachexp_school, data = CF)
summary(CFmr3)
```
```{r}
plot(CFmr3)
```

```{r}
CFmr2 <- lm(Zsagarin ~ Fravg + Soavg + Jravg + conference, data = CF)
summary(CFmr2)
```

I tried multiple things included power transformations but I ended up finding that the interaction terms seemed to benefit my R-squared value and lower my p-value more significantly. However, I ended up not adding an interaction term to the model due to it messing making the plots have some form of linearity to them. I played around with different interaction terms and from what I tested several had low p-values but most if not all made the plotss worse so I reverted back to my original model (CFmr2).

2)

```{r}
donner <- read_xlsx("donner.xlsx")
```

```{r}
train <- sample(45, 30, replace = FALSE)
train

traindata <- donner[train,]
traindata

testdata <- donner[-train,]
testdata
```

```{r}
training <- glm(Survived ~ Multifam + Sex + Age,family=binomial(link='logit'),data=traindata)
summary(training)
```

```{r}
probabilities <- predict(training, testdata, type = "response")
probabilities

prediction <- rep("Did Not Survive", 15)
prediction[probabilities > 0.5] <- "Survived"


table(prediction, testdata$Survived)  

```

This model was a decent predictor of whether the set of travelers survived or died. This model got 11 out of 15 correct giving us the correct answer 73% of the time.


```{r}
library(class)

trainingindex <- sample(nrow(donner), nrow(traindata), replace = FALSE)

train <- cbind(donner$Multifam, donner$Sex, donner$Age)[trainingindex, ]

test <- cbind(donner$Multifam, donner$Sex, donner$Age)[-trainingindex, ]

class <- donner$Survived[trainingindex]

knnpred <- knn(train, test, class, k = 1)

knnpred <- factor(ifelse(knnpred == 1, "Survived", "Died"))
table(knnpred, testdata$Survived)
```
Based on the above table the K Nearest Neighbors approach to this data gives us a worse percentage than the logistic regression. The K Nearest Neighbors correctly predicted 9 out of the 15 total tests giving us a percentage of 60%. This is lower than the logistic regression percentage of 73% by a significant percentage. In this case the logistic regression would be better to predict whether the members of the Donner Party survived or not on their travels.
